Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-19 sections/01-introduction.tex
Update on Overleaf.

Measurement aims to obtain accurate and reliable data about the current state of a system. 
A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
When introducing new measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods \parencite{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series \parencite{watson2010method}. 
Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value \parencite{bland1986statistical}. 
Trending considerations, that is, the consistency of the test methods development with the gold standard, is a field of active research in the last years \parencite{Saugel2015,saugel2018error,hiraishi2021concordance}. 
In this paper, we build upon the existing methods and extend them with new functionality.

As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
However, regardless of the application area, most conventional method comparisons and assessments consider local absolute deviations without assessing whether the correct direction of change is being predicted or measured. 
In this paper, we develop trending evaluation methods that complement current evaluation measures with trending measurements.

@@

The remainder of this paper is structured as follows. 
Section \ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts. 
In particular, we introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries. 
In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting. 
We conclude the paper in Section~\ref{sec:conclusion}.

>>>>

Measurement is more technical and aims to obtain accurate and reliable data about the current state of a system based on sensors. 
A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods \parencite{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series \parencite{watson2010method}. 
Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \parencite{bland1986statistical}. 
Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years \parencite{Saugel2015,saugel2018error,hiraishi2021concordance}.

As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
We apply the existing trending methods from measurement evaluation to the other fields and extend their functionality.
In this paper, we develop trending evaluation methods that complement the current evaluation with trending measurements.

@@

The remainder of this paper is structured as follows. 
Section \ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts. 
We introduce a new graphical method for evaluating trending, the conditional trending plot, which can simultaneously assess different intervals and asymmetries. 
In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting. 
We conclude the paper in Section~\ref{sec:conclusion}.

### 2024-03-19 sections/03-trending.tex
Update on Overleaf.

As discussed in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by another measurement, a nowcast, or a forecast.
In this section, we want to specify trending and define different methods of measuring and visualizing trending.
Development always has respect to a time horizon, that is, the amount of time that has passed. 
In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the question at hand.
Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
Before introducing measures to evaluate trending, we remark on calculating the predicted change and notation in Section~\ref{subsec:notation}.
Section~\ref{subsec:trending-four-quadrant-plot} introduces trending basics and visualizes trending in a synthetic example.
In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise}, and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment.
Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:trending-probabilistic}.

@@

The definition of the \textit{predicted} change depends on the context.
Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$. 
We call $t$ the \textit{forecast time} and $\tau$ the \textit{issue time}.
We compute the predicted change by
\begin{equation}
  \diffxl =
\begin{cases}
(x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
(x_{t|t} - y_{t-l})^T_{t=l} & \text{otherwise}.
\end{cases} \label{eq:diffxl_nowcasting}
\end{equation}
When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet. 
Thus, the predicted change can be computed with the knowledge of time $t$, and the trending assessment can be used in practice at time $t$.

The notation is similar for forecasting: Let $x_{t | \tau}$ denote the forecast for forecast time $t$ and issue time $\tau$. 
The predicted change is computed by 
\begin{equation} \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l} \label{eq:diffxl_forecasting}
\end{equation}
The indices are chosen to be consistent with $\diffyl$. 
If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification as in Equation~\eqref{eq:diffxl_nowcasting} can be made.

The distinction between forecast and issue time is unnecessary in measurement analysis, as the measurement is typically available with a very short time lag. 
Thus, $x_t$ denotes the test method measurement for time $t$. 
The computation
\begin{equation}
  (x_{t} - x_{t-l})_{t=l}^T \label{eq:diffxl_measure}
\end{equation}
yields the change by the test method. 
It is computed purely by the test method without the gold standard $y_t$ to analyze whether the gold standard and test method changes are consistent.

>>>>

As discussed in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by another measurement, a nowcast, or a forecast.
In this section, we want to specify trending and define different methods of measuring and visualizing trending.
The development in trending assessment always has respect to a time horizon, that is, the amount of time that has passed. 
In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the question at hand.
Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
Before introducing measures to evaluate trending, we remark on calculating the predicted change and notation in Section~\ref{subsec:notation}.
Section~\ref{subsec:trending-four-quadrant-plot} introduces trending basics and visualizes trending in a synthetic example.
In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise}, and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment.
Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:trending-probabilistic}.

@@

The definition of the \textit{predicted} change depends on the context;
Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$. 
We call $t$ the \textit{forecast time} and $\tau$ the \textit{issue time}.
The predicted change is computed by
\begin{equation}
  \diffxl =
\begin{cases}
(x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
(x_{t|t} - y_{t-l})^T_{t=l} & \text{otherwise}.
\end{cases} \label{eq:diffxl_nowcasting}
\end{equation}
When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet. 
Thus, the predicted change can be computed with the knowledge of time $t$. Therefore, the trending information of the following section can be used in practice at time $t$.

The notation is similar for forecasting: Let $x_{t | \tau}$ denote the forecast for forecast time $t$ and issue time $\tau$. 
The predicted change is computed by 
\begin{equation} \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l}\label{eq:diffxl_forecasting}
\end{equation}
with the same structure as in the nowcasting case. 
The indices are chosen to be consistent with $\diffyl$. 
If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification as in Equation~\eqref{eq:diffxl_nowcasting} can be made.

The distinction between forecast and issue time is unnecessary in measurement analysis, as the measurement is typically available with a very short time lag. 
Thus, $x_t$ denotes the test method measurement for time $t$. 
The computation
\begin{equation}
  (x_{t} - x_{t-l})_{t=l}^T \label{eq:diffxl_measure}
\end{equation}
yields the change by the test method. 
It is computed purely by the test method without the gold standard $y_t$ to analyze whether the gold standard and test method changes are consistent within themselves.
